\documentclass[11pt]{article}

\usepackage[a4paper,margin=1in]{geometry}
\usepackage{amsmath,amssymb,mathtools}
\usepackage{booktabs}
\usepackage{hyperref}

\newcommand{\dd}{\mathrm d}
\newcommand{\ii}{\mathrm i}
\newcommand{\eps}{\epsilon}
\newcommand{\U}{\mathcal U}
\newcommand{\F}{\mathcal F}
\newcommand{\hU}{\widehat{\mathcal U}}
\newcommand{\hF}{\widehat{\mathcal F}}
\newcommand{\Dlt}{\Delta}

\title{Fast Sector Decomposition from Black-Box Symanzik Polynomials}
\author{}
\date{}

\begin{document}

\maketitle

\section{The example triangle}

We consider the scalar one-loop triangle with one incoming off-shell momentum
and two outgoing massless momenta,
\begin{equation}
  p_0=p_1+p_2,
  \qquad
  p_1^2=p_2^2=0,
  \qquad
  p_0^2=s=m_0^2 .
\end{equation}
All three internal propagators are taken to have the same mass \(m\).  A
convenient routing is
\begin{align}
  D_0 &= \ell^2-m^2+\ii0,\\
  D_1 &= (\ell+p_1)^2-m^2+\ii0,\\
  D_2 &= (\ell-p_2)^2-m^2+\ii0 .
\end{align}
The scalar integral is
\begin{equation}
  C_0(s;m^2)
  =
  \int \frac{\dd^D\ell}{\ii\pi^{D/2}}\,
  \frac{1}{D_0D_1D_2}.
\end{equation}
With Feynman parameters \(x_i\geq 0\), the projective representation is
\begin{equation}
  C_0(s;m^2)
  =
  -\Gamma(\alpha)
  \int_0^\infty \dd x_0\,\dd x_1\,\dd x_2\,
  \delta(1-x_0-x_1-x_2)\,
  \frac{\U^{3-D}}{\F^\alpha},
  \qquad
  \alpha=3-\frac{D}{2}.
\end{equation}
For this routing,
\begin{equation}
  \U=x_0+x_1+x_2,
\end{equation}
and
\begin{equation}
  \F
  =
  m^2\U^2
  -
  s\,x_1x_2
  -
  \ii0 .
\end{equation}
Solving the simplex constraint by
\begin{equation}
  x_0=1-x_1-x_2
\end{equation}
gives
\begin{equation}
  0\leq x_1\leq 1,
  \qquad
  0\leq x_2\leq 1-x_1,
\end{equation}
and therefore
\begin{equation}
  \U=1,
  \qquad
  \F=m^2-s\,x_1x_2-\ii0 .
\end{equation}
For \(D=4-2\eps\), one has \(\alpha=1+\eps\), and
\begin{equation}
  C_0(s;m^2)
  =
  -\Gamma(1+\eps)
  \int_0^1 \dd x_1
  \int_0^{1-x_1} \dd x_2\,
  \left(m^2-s\,x_1x_2-\ii0\right)^{-1-\eps}.
\end{equation}
Throughout the rest of the note, sector integrands \(I_s\) do not include the
global prefactor \(-\Gamma(\alpha)\).  Thus an exact decomposition has the form
\begin{equation}
  C_0(s;m^2)
  =
  -\Gamma(\alpha)
  \sum_s
  \int_{\mathcal C_s}\dd y_s\,
  I_s(y_s).
\end{equation}

\subsection{Endpoint sectors}

The primary endpoint decomposition splits the simplex into two sectors.

\paragraph{Sector \(S_1\), \(x_1\geq x_2\).}
\begin{equation}
  X_{S_1}(t,z)
  =
  (x_0,x_1,x_2)
  =
  \left(
    1-t,\,
    \frac{t}{1+z},\,
    \frac{tz}{1+z}
  \right),
  \qquad
  0\leq t,z\leq 1 .
\end{equation}
The Jacobian is
\begin{equation}
  J_{S_1}(t,z)=\frac{t}{(1+z)^2}.
\end{equation}
The pulled-back second Symanzik polynomial is
\begin{equation}
  \F_{S_1}(t,z)
  =
  m^2
  -
  s\,\frac{t^2z}{(1+z)^2}
  -
  \ii0 .
\end{equation}
In the massless endpoint case \(m=0\),
\begin{equation}
  \F_{S_1}(t,z)
  =
  (-s-\ii0)\,\frac{t^2z}{(1+z)^2}.
\end{equation}
The expected monomial is therefore
\begin{equation}
  M_{S_1}(t,z)=t^2z .
\end{equation}

\paragraph{Sector \(S_2\), \(x_2\geq x_1\).}
\begin{equation}
  X_{S_2}(t,z)
  =
  (x_0,x_1,x_2)
  =
  \left(
    1-t,\,
    \frac{tz}{1+z},\,
    \frac{t}{1+z}
  \right),
  \qquad
  0\leq t,z\leq 1 .
\end{equation}
The Jacobian is again
\begin{equation}
  J_{S_2}(t,z)=\frac{t}{(1+z)^2}.
\end{equation}
The pulled-back polynomial is
\begin{equation}
  \F_{S_2}(t,z)
  =
  m^2
  -
  s\,\frac{t^2z}{(1+z)^2}
  -
  \ii0 .
\end{equation}
In the massless endpoint case,
\begin{equation}
  \F_{S_2}(t,z)
  =
  (-s-\ii0)\,\frac{t^2z}{(1+z)^2},
  \qquad
  M_{S_2}(t,z)=t^2z .
\end{equation}
With these definitions, the two-sector representation of the original
triangle is
\begin{align}
  C_0(s;m^2)
  &=
  -\Gamma(\alpha)
  \sum_{a=1}^{2}
  \int_0^1\dd t
  \int_0^1\dd z\,
  I_{S_a}(t,z),
  \\
  I_{S_a}(t,z)
  &=
  \frac{t}{(1+z)^2}\,
  \U(X_{S_a}(t,z))^{3-D}\,
  \F(X_{S_a}(t,z))^{-\alpha}.
\end{align}
Equivalently, after inserting the explicit polynomial,
\begin{equation}
  C_0(s;m^2)
  =
  -\Gamma(\alpha)
  \sum_{a=1}^{2}
  \int_0^1\dd t
  \int_0^1\dd z\,
  \frac{t}{(1+z)^2}
  \left(
    m^2
    -
    s\,\frac{t^2z}{(1+z)^2}
    -
    \ii0
  \right)^{-\alpha}.
\end{equation}
For \(D=4-2\eps\) and \(m=0\), this becomes
\begin{equation}
  C_0(s;0)
  =
  -\Gamma(1+\eps)
  \sum_{a=1}^{2}
  \int_0^1\dd t
  \int_0^1\dd z\,
  t^{-1-2\eps}
  z^{-1-\eps}
  G_{S_a}(t,z),
\end{equation}
where
\begin{equation}
  G_{S_a}(t,z)
  =
  (-s-\ii0)^{-1-\eps}(1+z)^{2\eps}.
\end{equation}

\subsection{Massive threshold sectors}

For \(m\neq 0\), the endpoint singularity is lifted.  A threshold singularity
appears at
\begin{equation}
  s=4m^2,
\end{equation}
at the point
\begin{equation}
  x_0=0,
  \qquad
  x_1=x_2=\frac12 .
\end{equation}
Write
\begin{equation}
  x_0=a,
  \qquad
  x_1=\frac{1-a+d}{2},
  \qquad
  x_2=\frac{1-a-d}{2},
\end{equation}
and split
\begin{equation}
  d=\sigma(1-a)b,
  \qquad
  \sigma=\pm1,
  \qquad
  0\leq b\leq 1 .
\end{equation}
At threshold,
\begin{equation}
  \F
  =
  m^2\left[a(2-a)+(1-a)^2b^2\right]-\ii0 .
\end{equation}
The overlap \(a\sim b^2\) is blown up by two sectors for each sign
\(\sigma\).

\paragraph{Sector \(T_{\sigma,1}\), \(b^2\leq a\).}
Use \(a=r^2\), \(b=ry\).  Then
\begin{equation}
  X_{T_{\sigma,1}}(r,y)
  =
  \left(
    r^2,\,
    \frac{(1-r^2)(1+\sigma ry)}{2},\,
    \frac{(1-r^2)(1-\sigma ry)}{2}
  \right),
  \qquad
  0\leq r,y\leq 1 .
\end{equation}
The Jacobian is
\begin{equation}
  J_{T_{\sigma,1}}(r,y)=r^2(1-r^2).
\end{equation}
The polynomial factorizes as
\begin{equation}
  \F_{T_{\sigma,1}}(r,y)
  =
  m^2r^2
  \left[
    (2-r^2)+(1-r^2)^2y^2
  \right]
  -
  \ii0 .
\end{equation}
The expected monomial is
\begin{equation}
  M_{T_{\sigma,1}}(r,y)=r^2 .
\end{equation}

\paragraph{Sector \(T_{\sigma,2}\), \(a\leq b^2\).}
Use \(a=b^2y\).  Then
\begin{equation}
  X_{T_{\sigma,2}}(b,y)
  =
  \left(
    b^2y,\,
    \frac{(1-b^2y)(1+\sigma b)}{2},\,
    \frac{(1-b^2y)(1-\sigma b)}{2}
  \right),
  \qquad
  0\leq b,y\leq 1 .
\end{equation}
The Jacobian is
\begin{equation}
  J_{T_{\sigma,2}}(b,y)
  =
  \frac12\,b^2(1-b^2y).
\end{equation}
The polynomial factorizes as
\begin{equation}
  \F_{T_{\sigma,2}}(b,y)
  =
  m^2b^2
  \left[
    y(2-b^2y)+(1-b^2y)^2
  \right]
  -
  \ii0 .
\end{equation}
The expected monomial is
\begin{equation}
  M_{T_{\sigma,2}}(b,y)=b^2 .
\end{equation}
At exact threshold, the four threshold sectors reconstruct the original
massive triangle as
\begin{align}
  C_0(4m^2;m^2)
  &=
  -\Gamma(\alpha)
  \sum_{\sigma=\pm1}
  \left[
    \int_0^1\dd r
    \int_0^1\dd y\,
    I_{T_{\sigma,1}}(r,y)
  \right.
  \notag\\
  &\hspace{4.7cm}
  \left.
    +
    \int_0^1\dd b
    \int_0^1\dd y\,
    I_{T_{\sigma,2}}(b,y)
  \right],
  \\
  I_{T_{\sigma,1}}(r,y)
  &=
  r^2(1-r^2)\,
  \F(X_{T_{\sigma,1}}(r,y))^{-\alpha},
  \\
  I_{T_{\sigma,2}}(b,y)
  &=
  \frac12 b^2(1-b^2y)\,
  \F(X_{T_{\sigma,2}}(b,y))^{-\alpha}.
\end{align}
Here \(\U=1\) after the simplex constraint, so no separate power of \(\U\)
appears in the last two lines.

\section{Direct approach}

In the direct approach, each sector map \(X_s\) is substituted explicitly into
\(\U\) and \(\F\).  If the pulled-back polynomial factorizes as
\begin{equation}
  \F_s(y)
  =
  \F(X_s(y))
  =
  M_s(y)\,\phi_s(y),
\end{equation}
then the sector integrand is built as
\begin{equation}
  I_s^{\mathrm{dir}}(y)
  =
  J_s(y)\,
  \U(X_s(y))^{3-D}\,
  M_s(y)^{-\alpha}\,
  \phi_s(y)^{-\alpha}.
\end{equation}
The corresponding contribution to the original scalar integral is always
understood as
\begin{equation}
  C_0
  =
  -\Gamma(\alpha)
  \sum_s
  \int_{\mathcal C_s}\dd y_s\,
  I_s^{\mathrm{dir}}(y_s),
\end{equation}
with the sector domains \(\mathcal C_s\) specified explicitly below.

\subsection{\texorpdfstring{Endpoint sector \(S_1\)}{Endpoint sector S1}}

For \(m=0\),
\begin{equation}
  \U_{S_1}(t,z)=1,
\end{equation}
and
\begin{equation}
  \F_{S_1}(t,z)
  =
  (-s-\ii0)\,\frac{t^2z}{(1+z)^2}
  =
  M_{S_1}(t,z)\,\phi_{S_1}(t,z),
\end{equation}
with
\begin{equation}
  M_{S_1}(t,z)=t^2z,
  \qquad
  \phi_{S_1}(t,z)=\frac{-s-\ii0}{(1+z)^2}.
\end{equation}
Therefore
\begin{align}
  I_{S_1}^{\mathrm{dir}}(t,z)
  &=
  \frac{t}{(1+z)^2}\,
  (t^2z)^{-\alpha}
  \left(\frac{-s-\ii0}{(1+z)^2}\right)^{-\alpha}.
\end{align}
For \(D=4-2\eps\), i.e. \(\alpha=1+\eps\),
\begin{equation}
  I_{S_1}^{\mathrm{dir}}(t,z)
  =
  (-s-\ii0)^{-1-\eps}\,
  t^{-1-2\eps}\,
  z^{-1-\eps}\,
  (1+z)^{2\eps}.
\end{equation}

\subsection{\texorpdfstring{Endpoint sector \(S_2\)}{Endpoint sector S2}}

By symmetry,
\begin{equation}
  \F_{S_2}(t,z)
  =
  (-s-\ii0)\,\frac{t^2z}{(1+z)^2},
  \qquad
  M_{S_2}(t,z)=t^2z,
\end{equation}
and hence
\begin{equation}
  I_{S_2}^{\mathrm{dir}}(t,z)
  =
  (-s-\ii0)^{-1-\eps}\,
  t^{-1-2\eps}\,
  z^{-1-\eps}\,
  (1+z)^{2\eps}.
\end{equation}
The full massless triangle is therefore
\begin{align}
  C_0(s;0)
  &=
  -\Gamma(1+\eps)
  \sum_{a=1}^{2}
  \int_0^1\dd t
  \int_0^1\dd z\,
  I_{S_a}^{\mathrm{dir}}(t,z)
  \notag\\
  &=
  -2\Gamma(1+\eps)
  (-s-\ii0)^{-1-\eps}
  \int_0^1\dd t
  \int_0^1\dd z\,
  t^{-1-2\eps}
  z^{-1-\eps}
  (1+z)^{2\eps}.
\end{align}

\subsection{Threshold sectors}

At \(s=4m^2\), the direct threshold integrands are
\begin{align}
  I_{T_{\sigma,1}}^{\mathrm{dir}}(r,y)
  &=
  r^2(1-r^2)\,
  (r^2)^{-\alpha}
  \left(
    m^2\left[(2-r^2)+(1-r^2)^2y^2\right]-\ii0
  \right)^{-\alpha},
  \\
  I_{T_{\sigma,2}}^{\mathrm{dir}}(b,y)
  &=
  \frac12 b^2(1-b^2y)\,
  (b^2)^{-\alpha}
  \left(
    m^2\left[y(2-b^2y)+(1-b^2y)^2\right]-\ii0
  \right)^{-\alpha}.
\end{align}
The dependence on \(\sigma\) is only in the map \(X_T\), not in the
factorized polynomial for this symmetric equal-mass triangle.
These four direct threshold sectors reconstruct the threshold integral as
\begin{align}
  C_0(4m^2;m^2)
  &=
  -\Gamma(\alpha)
  \sum_{\sigma=\pm1}
  \left[
    \int_0^1\dd r
    \int_0^1\dd y\,
    I_{T_{\sigma,1}}^{\mathrm{dir}}(r,y)
  \right.
  \notag\\
  &\hspace{4.8cm}
  \left.
    +
    \int_0^1\dd b
    \int_0^1\dd y\,
    I_{T_{\sigma,2}}^{\mathrm{dir}}(b,y)
  \right].
\end{align}

\section{Indirect approach}

The indirect approach keeps the Symanzik polynomials closed.  We assume only
black-box evaluators
\begin{equation}
  \hU(x_0,x_1,x_2),
  \qquad
  \hF(x_0,x_1,x_2),
\end{equation}
together with their partial derivatives with respect to the original
parameters \(x_i\).  In practice, these derivative jets can be computed
efficiently with dual numbers.

For a sector map \(X_s\), define the pullbacks
\begin{equation}
  U_s(y)=\hU(X_s(y)),
  \qquad
  F_s(y)=\hF(X_s(y)).
\end{equation}
Derivatives with respect to sector variables are obtained by chain rules.  For
example,
\begin{equation}
  \partial_a F_s(y)
  =
  \sum_i
  \frac{\partial \hF}{\partial x_i}(X_s(y))\,
  \frac{\partial X_{s,i}}{\partial y_a}(y),
\end{equation}
and
\begin{align}
  \partial_a\partial_b F_s(y)
  &=
  \sum_{i,j}
  \frac{\partial^2 \hF}{\partial x_i\partial x_j}(X_s(y))\,
  \frac{\partial X_{s,i}}{\partial y_a}(y)\,
  \frac{\partial X_{s,j}}{\partial y_b}(y)
  \notag\\
  &\quad
  +
  \sum_i
  \frac{\partial \hF}{\partial x_i}(X_s(y))\,
  \frac{\partial^2 X_{s,i}}{\partial y_a\partial y_b}(y).
\end{align}
Higher derivatives are obtained similarly, or more directly by propagating
dual-number jets through the known sector map and then through the black-box
polynomial evaluator.

The object to construct is the regular coefficient \(\phi_s\) defined by
\begin{equation}
  F_s(y)=M_s(y)\,\phi_s(y),
\end{equation}
where \(M_s\) is the expected monomial for that sector.  Once \(\phi_s\) is
available, the sector integrand is
\begin{equation}
  I_s^{\mathrm{ind}}(y)
  =
  J_s(y)\,
  U_s(y)^{3-D}\,
  M_s(y)^{-\alpha}\,
  \phi_s(y)^{-\alpha}.
\end{equation}
At interior points this is pointwise identical to direct substitution, since
\begin{equation}
  \phi_s(y)=\frac{\hF(X_s(y))}{M_s(y)}
\end{equation}
there.
For the endpoint decomposition before any monomial extraction, the black-box
form of the exact triangle is
\begin{align}
  C_0(s;m^2)
  &=
  -\Gamma(\alpha)
  \sum_{a=1}^{2}
  \int_0^1\dd t
  \int_0^1\dd z\,
  \frac{t}{(1+z)^2}
  \hU(X_{S_a}(t,z))^{3-D}
  \hF(X_{S_a}(t,z))^{-\alpha}.
\end{align}
When \(m=0\), and hence
\(\hF(X_{S_a}(t,z))=t^2z\,\phi_{S_a}(t,z)\), this same equality becomes
\begin{align}
  C_0(s;0)
  &=
  -\Gamma(1+\eps)
  \sum_{a=1}^{2}
  \int_0^1\dd t
  \int_0^1\dd z\,
  \frac{t}{(1+z)^2}\,
  \hU(X_{S_a}(t,z))^{-1+2\eps}
  \notag\\
  &\hspace{3.2cm}\times
  (t^2z)^{-1-\eps}\,
  \phi_{S_a}(t,z)^{-1-\eps}.
\end{align}

\subsection{Using Cauchy remainder formula}

Suppose
\begin{equation}
  M_s(y)=\prod_{a=1}^n y_a^{r_a},
  \qquad
  r_a\in\mathbb N .
\end{equation}
If \(F_s\) is divisible by \(M_s\), then repeated use of the Cauchy form of
the Taylor remainder gives
\begin{equation}
  \phi_s(y)
  =
  \prod_{a:r_a>0}
  \left[
    \frac{1}{(r_a-1)!}
    \int_0^1 \dd\lambda_a\,
    (1-\lambda_a)^{r_a-1}
  \right]
  \left(
    \prod_{a:r_a>0}
    \partial_{y_a}^{r_a}
  \right)
  F_s(y^{(\lambda)}),
\end{equation}
where
\begin{equation}
  y_a^{(\lambda)}=
  \begin{cases}
    \lambda_a y_a, & r_a>0,\\
    y_a, & r_a=0 .
  \end{cases}
\end{equation}
This formula does not require opening \(\hF\).  The derivatives of
\(F_s=\hF\circ X_s\) are supplied by the chain rules above.

For the first endpoint sector,
\begin{equation}
  M_{S_1}(t,z)=t^2z.
\end{equation}
Introduce dummy variables \(T,Z\) and write
\begin{equation}
  F_{S_1}(T,Z)
  =
  \hF\left(
    1-T,\,
    \frac{T}{1+Z},\,
    \frac{TZ}{1+Z}
  \right).
\end{equation}
Then
\begin{equation}
  \phi_{S_1}^{\mathrm{C}}(t,z)
  =
  \int_0^1 \dd\lambda\,(1-\lambda)
  \int_0^1 \dd\mu\,
  \left[
    \partial_T^2\partial_Z F_{S_1}(T,Z)
  \right]_{T=\lambda t,\;Z=\mu z}.
\end{equation}
Equivalently, keeping the black box unopened,
\begin{align}
  \phi_{S_1}^{\mathrm{C}}(t,z)
  &=
  \int_0^1 \dd\lambda\,(1-\lambda)
  \int_0^1 \dd\mu\,
  \partial_T^2\partial_Z
  \hF\left(
    1-T,\,
    \frac{T}{1+Z},\,
    \frac{TZ}{1+Z}
  \right)
  \bigg|_{T=\lambda t,\;Z=\mu z}.
\end{align}
The second endpoint sector is obtained with
\begin{equation}
  F_{S_2}(T,Z)
  =
  \hF\left(
    1-T,\,
    \frac{TZ}{1+Z},\,
    \frac{T}{1+Z}
  \right),
\end{equation}
and the same Cauchy formula,
\begin{equation}
  \phi_{S_2}^{\mathrm{C}}(t,z)
  =
  \int_0^1 \dd\lambda\,(1-\lambda)
  \int_0^1 \dd\mu\,
  \left[
    \partial_T^2\partial_Z F_{S_2}(T,Z)
  \right]_{T=\lambda t,\;Z=\mu z}.
\end{equation}
Using these Cauchy-remainder coefficients, the complete massless triangle is
\begin{align}
  C_0(s;0)
  &=
  -\Gamma(1+\eps)
  \sum_{a=1}^{2}
  \int_0^1\dd t
  \int_0^1\dd z\,
  \frac{t}{(1+z)^2}\,
  \hU(X_{S_a}(t,z))^{-1+2\eps}
  \notag\\
  &\hspace{3.2cm}\times
  (t^2z)^{-1-\eps}\,
  \left[\phi_{S_a}^{\mathrm{C}}(t,z)\right]^{-1-\eps}.
\end{align}

If we momentarily open the explicit massless triangle polynomial, then
\begin{equation}
  F_{S_1}(T,Z)
  =
  (-s-\ii0)\,\frac{T^2Z}{(1+Z)^2}.
\end{equation}
Therefore
\begin{equation}
  \partial_T^2\partial_Z F_{S_1}(T,Z)
  =
  2(-s-\ii0)\,\frac{1-Z}{(1+Z)^3}.
\end{equation}
The explicit \((\lambda,\mu)\)-integrand is
\begin{equation}
  (1-\lambda)\,
  2(-s-\ii0)\,
  \frac{1-\mu z}{(1+\mu z)^3}.
\end{equation}
Consequently,
\begin{align}
  \phi_{S_1}^{\mathrm{C}}(t,z)
  &=
  \int_0^1 \dd\lambda\,(1-\lambda)
  \int_0^1 \dd\mu\,
  2(-s-\ii0)\,
  \frac{1-\mu z}{(1+\mu z)^3}
  \notag\\
  &=
  \frac{-s-\ii0}{(1+z)^2},
\end{align}
as in the direct approach.
The same computation holds in sector \(S_2\).  Hence the explicit
Cauchy-remainder reconstruction of the massless triangle is
\begin{align}
  C_0(s;0)
  &=
  -\Gamma(1+\eps)
  \sum_{a=1}^{2}
  \int_0^1\dd t
  \int_0^1\dd z\,
  \frac{t}{(1+z)^2}
  (t^2z)^{-1-\eps}
  \left(\frac{-s-\ii0}{(1+z)^2}\right)^{-1-\eps}
  \notag\\
  &=
  -2\Gamma(1+\eps)
  (-s-\ii0)^{-1-\eps}
  \int_0^1\dd t
  \int_0^1\dd z\,
  t^{-1-2\eps}
  z^{-1-\eps}
  (1+z)^{2\eps}.
\end{align}

The auxiliary integrals are well behaved because the assumed monomial
divisibility implies that \(F_s=M_s\phi_s\) with \(\phi_s\) regular at the
sector boundaries.  The Cauchy-remainder integrands are derivatives of a
regular pullback evaluated on a compact domain.  In this explicit sector,
\begin{equation}
  \frac{1-\mu z}{(1+\mu z)^3}
\end{equation}
is bounded for \(0\leq \mu,z\leq 1\).

\subsection{Using subtraction terms}

The alternative is to avoid the auxiliary Cauchy-remainder integrals.  At
interior points one simply evaluates
\begin{equation}
  \phi_s(y)=\frac{\hF(X_s(y))}{M_s(y)}.
\end{equation}
Boundary values are supplied by the continuous extension, using derivatives.
For the first endpoint sector,
\begin{align}
  \phi_{S_1}(0,z)
  &=
  \frac{1}{2z}
  \left[
    \partial_T^2 F_{S_1}(T,z)
  \right]_{T=0},
  \\
  \phi_{S_1}(t,0)
  &=
  \frac{1}{t^2}
  \left[
    \partial_Z F_{S_1}(t,Z)
  \right]_{Z=0},
  \\
  \phi_{S_1}(0,0)
  &=
  \frac12
  \left[
    \partial_T^2\partial_Z F_{S_1}(T,Z)
  \right]_{T=0,\;Z=0}.
\end{align}
Again, the derivatives of \(F_{S_1}\) are not symbolic derivatives of an opened
polynomial.  They are derivatives of \(\hF\circ X_{S_1}\), obtained from the
black-box derivative jets and the chain rule.

For \(D=4-2\eps\), write each factorized endpoint integrand as
\begin{equation}
  I_{S_a}^{\mathrm{ind}}(t,z)
  =
  t^{-1-2\eps}\,
  z^{-1-\eps}\,
  G_{S_a}(t,z),
  \qquad
  a=1,2,
\end{equation}
where
\begin{equation}
  G_{S_a}(t,z)
  =
  \frac{1}{(1+z)^2}\,
  U_{S_a}(t,z)^{-1+2\eps}\,
  \phi_{S_a}(t,z)^{-1-\eps}.
\end{equation}
The two-dimensional endpoint subtraction identity is
\begin{align}
  &\int_0^1 \dd t\int_0^1 \dd z\,
  t^{-1-2\eps}z^{-1-\eps}G(t,z)
  \notag\\
  &=
  \int_0^1 \dd t\int_0^1 \dd z\,
  t^{-1-2\eps}z^{-1-\eps}
  \left[
    G(t,z)-G(0,z)-G(t,0)+G(0,0)
  \right]
  \notag\\
  &\quad
  -\frac{1}{2\eps}
  \int_0^1 \dd z\,
  z^{-1-\eps}
  \left[
    G(0,z)-G(0,0)
  \right]
  \notag\\
  &\quad
  -\frac{1}{\eps}
  \int_0^1 \dd t\,
  t^{-1-2\eps}
  \left[
    G(t,0)-G(0,0)
  \right]
  +
  \frac{G(0,0)}{2\eps^2}.
\end{align}
Each bracket is less singular at the corresponding endpoint.  This is the
subtraction-term alternative to constructing \(\phi_s\) with auxiliary
integrals.
Applied sector by sector, it gives the full massless triangle as
\begin{align}
  C_0(s;0)
  &=
  -\Gamma(1+\eps)
  \sum_{a=1}^{2}
  \Bigg\{
  \int_0^1 \dd t\int_0^1 \dd z\,
  t^{-1-2\eps}z^{-1-\eps}
  \Big[
    G_{S_a}(t,z)-G_{S_a}(0,z)
  \notag\\
  &\hspace{5.1cm}
    -G_{S_a}(t,0)+G_{S_a}(0,0)
  \Big]
  \notag\\
  &\qquad
  -\frac{1}{2\eps}
  \int_0^1 \dd z\,
  z^{-1-\eps}
  \left[
    G_{S_a}(0,z)-G_{S_a}(0,0)
  \right]
  \notag\\
  &\qquad
  -\frac{1}{\eps}
  \int_0^1 \dd t\,
  t^{-1-2\eps}
  \left[
    G_{S_a}(t,0)-G_{S_a}(0,0)
  \right]
  +
  \frac{G_{S_a}(0,0)}{2\eps^2}
  \Bigg\}.
\end{align}

For the explicit massless triangle in sector \(S_1\),
\begin{equation}
  U_{S_1}(t,z)=1,
  \qquad
  \phi_{S_1}(t,z)=\frac{-s-\ii0}{(1+z)^2}.
\end{equation}
Thus
\begin{equation}
  G_{S_1}^{\mathrm{exp}}(t,z)
  =
  (-s-\ii0)^{-1-\eps}(1+z)^{2\eps}.
\end{equation}
This function is independent of \(t\).  Therefore
\begin{equation}
  G_{S_1}^{\mathrm{exp}}(t,z)
  -
  G_{S_1}^{\mathrm{exp}}(0,z)
  -
  G_{S_1}^{\mathrm{exp}}(t,0)
  +
  G_{S_1}^{\mathrm{exp}}(0,0)
  =
  0,
\end{equation}
and
\begin{equation}
  G_{S_1}^{\mathrm{exp}}(t,0)
  -
  G_{S_1}^{\mathrm{exp}}(0,0)
  =
  0.
\end{equation}
The only non-trivial subtraction piece is
\begin{equation}
  G_{S_1}^{\mathrm{exp}}(0,z)
  -
  G_{S_1}^{\mathrm{exp}}(0,0)
  =
  (-s-\ii0)^{-1-\eps}
  \left[
    (1+z)^{2\eps}-1
  \right].
\end{equation}
The explicitly subtracted result for one endpoint sector is therefore
\begin{align}
  &\int_0^1 \dd t\int_0^1 \dd z\,
  I_{S_1}^{\mathrm{dir}}(t,z)
  \notag\\
  &=
  -\frac{(-s-\ii0)^{-1-\eps}}{2\eps}
  \int_0^1 \dd z\,
  z^{-1-\eps}
  \left[
    (1+z)^{2\eps}-1
  \right]
  +
  \frac{(-s-\ii0)^{-1-\eps}}{2\eps^2}.
\end{align}
The second sector gives the same contribution, so the full massless triangle is
\begin{align}
  C_0(s;0)
  &=
  -2\Gamma(1+\eps)
  \left\{
  -\frac{(-s-\ii0)^{-1-\eps}}{2\eps}
  \int_0^1 \dd z\,
  z^{-1-\eps}
  \left[
    (1+z)^{2\eps}-1
  \right]
  +
  \frac{(-s-\ii0)^{-1-\eps}}{2\eps^2}
  \right\}.
\end{align}
This expression was obtained without requiring any symbolic expansion in an
implementation: only \(X_{S_a}\), \(J_{S_a}\), the known monomial \(t^2z\), and
black-box values and derivative limits of \(\hU\) and \(\hF\) are required in
each endpoint sector.

\section{Python prototype in \texttt{FSD}}

The folder \texttt{FSD} contains a first implementation of the construction
above.  The executable entry point is
\begin{equation}
  \texttt{FSD/FSD.py}.
\end{equation}
It uses a local Python virtual environment,
\begin{equation}
  \texttt{FSD/.venv},
\end{equation}
and a local checkout of OneLOopBridge,
\begin{equation}
  \texttt{FSD/external/OneLOopBridge}.
\end{equation}
The benchmark value is obtained from the Python binding
\begin{equation}
  \texttt{oneloop\_bridge.three\_point(0,0,s,m2,m2,m2)}.
\end{equation}

\subsection{Separation of prepared sectors and processing}

The code mirrors the split used in this note.  First, endpoint sectors are
prepared independently of the Symanzik polynomials.  Each sector stores only
the data
\begin{equation}
  \left\{
    X_s,\,
    J_s,\,
    M_s
  \right\}.
\end{equation}
For the current prototype these are just \(S_1\) and \(S_2\).  Threshold
sectors are intentionally not implemented yet.

The processing layer then receives the prepared sectors together with two
black-box Symbolica evaluators,
\begin{equation}
  \hU(x_0,x_1,x_2;s,m^2),
  \qquad
  \hF(x_0,x_1,x_2;s,m^2).
\end{equation}
After these evaluators are built, the processing code does not inspect or
symbolically rewrite the expressions for \(\U\) or \(\F\).  In particular,
\(\phi_s\) is obtained from
\begin{equation}
  \phi_s(t,z)
  =
  \frac{\hF(X_s(t,z))}{M_s(t,z)}
\end{equation}
at interior points, and from Symbolica dual-number derivatives at boundary
points.  The dual-number jets are propagated through the known map \(X_s\) and
then through the black-box evaluator for \(\hF\).  This realizes the chain-rule
construction described in the previous sections.

\subsection{Havana sampling}

The prototype uses Symbolica's Havana integrator as a sampler rather than as a
single black-box scalar integrator.  One adaptive grid is used for all
endpoint-sector contributions:
\begin{equation}
  \text{discrete sector id}
  \quad\otimes\quad
  (t,z)\in[0,1]^2.
\end{equation}
Lower-support subtraction terms are localized into this same two-dimensional
domain by multiplying by a unit-density dummy integration over the unused
variable.  Thus the \(z\)-edge term is represented as a function of \((t,z)\)
which is independent of \(t\), and the corner term is represented as a constant
function of \((t,z)\).  This avoids maintaining separate Havana grids for
two-dimensional, one-dimensional, and zero-dimensional pieces.

For every batch, the code samples points from the single grid, evaluates all
Laurent coefficients on those same points, and fills its own accumulators.  The
Havana grid is trained only with the absolute value of the finite-part
integrand.  In Symbolica's sampled representation, the first entry of
\(\texttt{sample.weights}\) is the full Monte Carlo weight used in the manual
accumulator.

The integration loop reports progress with \(\texttt{progressbar2}\).  The bar
shows the current iteration, the accumulated number of sampled points against
the nominal target \(N_{\rm iter}N_{\rm batch}\), the live relative Monte Carlo
error maxed over the Laurent coefficients, and, when a benchmark is available,
the live pull maxed over the Laurent coefficients.  The live pull is computed
after applying the same Gamma and prefactor conventions used in the final
output table.

\subsection{Implemented modes and conventions}

The command-line interface supports two modes.  In massive mode, with
\(m>0\) and \(s<4m^2\), the integral is finite and the endpoint monomial
factorization is not applied.  The implementation evaluates
\begin{equation}
  C_0(s;m^2)
  =
  -
  \sum_{a=1}^{2}
  \int_0^1\dd t
  \int_0^1\dd z\,
  J_{S_a}(t,z)\,
  \hU(X_{S_a}(t,z))^{-1}\,
  \hF(X_{S_a}(t,z))^{-1}.
\end{equation}

In massless mode, the implementation uses the subtraction formula of
Section~3.2 and returns the coefficients of
\begin{equation}
  C_0(s;0)
  =
  \frac{C_{-2}}{\epsilon^2}
  +
  \frac{C_{-1}}{\epsilon}
  +
  C_0
  +
  \mathcal O(\epsilon).
\end{equation}
The default output convention is the OneLOopBridge convention.  In this
convention the universal Gamma/Euler factors are stripped from the displayed
Laurent coefficients.  For the massless endpoint representation this also
amounts to removing the finite convention term
\begin{equation}
  \frac{\pi^2}{6}\,C_{-2}
\end{equation}
from the direct sector-parametric finite coefficient.  With this convention,
the prototype compares directly to the values returned by OneLOopBridge.
The separate command-line option \(\texttt{--prefactor-convention}\) controls
only the overall scalar-integral normalization shown in the output table.  The
default value \(\texttt{raw}\) displays the raw OneLOopBridge normalization.
The value \(\texttt{feynman}\) multiplies both the FastSecDec coefficients and
the benchmark coefficients by
\begin{equation}
  \texttt{TO\_FEYNMAN}
  =
  -\frac{1}{16\pi^2},
\end{equation}
which is the OneLOopBridge factor converting raw scalar-integral coefficients
to its Feynman-normalized convention.  In either display convention, values
with Monte Carlo uncertainty are printed with the two-significant-digit
parenthesis notation.

\subsection{Implemented box example}

The prototype also contains a finite massive scalar box example with massless
external legs,
\begin{equation}
  D_0(0,0,0,0,s_{12},s_{23};m^2).
\end{equation}
The benchmark call is
\begin{equation}
  \texttt{oneloop\_bridge.four\_point(0,0,0,0,s12,s23,m2,m2,m2,m2)}.
\end{equation}
For equal internal masses, the simplex polynomial used by the implementation is
\begin{equation}
  \U=x_0+x_1+x_2+x_3,
  \qquad
  \F
  =
  m^2\U^2
  -
  s_{12}x_0x_2
  -
  s_{23}x_1x_3
  -
  \ii0.
\end{equation}
After imposing \(\U=1\), the finite four-point integral in \(D=4\) is
\begin{equation}
  D_0
  =
  \sum_{b=0}^{3}
  \int_0^1\dd y_0
  \int_0^1\dd y_1
  \int_0^1\dd y_2\,
  J_b(y)\,
  \hF(X_b(y))^{-2}.
\end{equation}
Here \(b\) labels the primary sector in which \(x_b\) is the largest Feynman
parameter.  In sector \(b\),
\begin{equation}
  x_b=\frac{1}{1+y_0+y_1+y_2},
  \qquad
  x_{j\neq b}=\frac{y_j}{1+y_0+y_1+y_2},
\end{equation}
with the three \(y_j\)'s assigned in increasing order to the non-dominant
parameters.  The Jacobian is
\begin{equation}
  J_b(y)=\frac{1}{(1+y_0+y_1+y_2)^4}.
\end{equation}
The same code path now also implements the massless endpoint-subtracted box.
For \(m=0\), each primary sector has the form
\begin{equation}
  \F(X_b(y))
  =
  \frac{A_b\,y_\ell+B_b\,y_r y_s}
       {(1+y_0+y_1+y_2)^2},
\end{equation}
where one of the two bilinears \(x_0x_2\) and \(x_1x_3\) contains the dominant
parameter and is therefore linear in the primary variables.  The implementation
does not need the symbolic values of \(A_b\) and \(B_b\); it only uses the known
positions of the linear variable \(y_\ell\) and product variables \(y_r,y_s\).
The three secondary sectors are
\begin{align}
  L:\quad& y_r=uv,\quad y_\ell=u,\quad y_s=w,
  &
  M_L&=u,
  \\
  P:\quad& y_r=u,\quad y_s=v,\quad y_\ell=uvw,
  &
  M_P&=uv,
  \\
  Q:\quad& y_r=u,\quad y_\ell=uv,\quad y_s=vw,
  &
  M_Q&=uv.
\end{align}
The secondary Jacobian equals the displayed monomial in each case, so the
regular prefactor after monomial extraction is again
\begin{equation}
  \frac{1}{(1+y_0(u,v,w)+y_1(u,v,w)+y_2(u,v,w))^4}.
\end{equation}
The residual coefficient
\begin{equation}
  \phi_{b\alpha}(u,v,w)
  =
  \frac{\F(X_{b\alpha}(u,v,w))}{M_\alpha(u,v,w)}
\end{equation}
is evaluated by the black-box \(\F\) away from the endpoint and by dual-number
Taylor coefficients at \(u=0\) and \(v=0\).  The one-monomial sector \(L\) uses
the one-variable subtraction formula, while the \(P\) and \(Q\) sectors use the
two-variable subtraction formula.

The raw coefficients produced directly by these massless box sectors are
converted to the OneLOopBridge stripped convention as
\begin{equation}
  D_{-2}=A_{-2},
  \qquad
  D_{-1}=A_{-1}+A_{-2},
  \qquad
  D_0=A_0+A_{-1}+\frac{\pi^2}{6}A_{-2}.
\end{equation}
The current massless box command validates Euclidean kinematics,
\(s_{12}<0\) and \(s_{23}<0\).  Timelike or threshold kinematics would require
a contour deformation or threshold regularization, which is outside the
present prototype.

\subsection{Example commands}

The massive below-threshold check can be run with
\begin{equation}
  \texttt{FSD/.venv/bin/python FSD/FSD.py --s 1.0 --m 1.0 --mode massive}.
\end{equation}
The same check in the Feynman-normalized prefactor convention can be displayed
with
\begin{align}
  &\texttt{FSD/.venv/bin/python FSD/FSD.py --s 1.0 --m 1.0 --mode massive}
  \notag\\
  &\texttt{--prefactor-convention feynman}.
\end{align}
The massless Euclidean check can be run with
\begin{equation}
  \texttt{FSD/.venv/bin/python FSD/FSD.py --s -1.0 --m 0.0 --mode massless}.
\end{equation}
The timelike command with \(s>0\) is intentionally rejected in the current
prototype because no contour deformation or threshold regularization is
implemented.
The massive box example can be run with
\begin{equation}
  \texttt{FSD/.venv/bin/python FSD/FSD.py --integral box --s12 0.5 --s23 0.7 --m 1.0}.
\end{equation}
The massless box check can be run with
\begin{equation}
  \texttt{FSD/.venv/bin/python FSD/FSD.py --integral box --s12 -1.0 --s23 -2.0 --m 0.0}.
\end{equation}
The corresponding timelike massless box command is rejected for the same
reason as the timelike triangle.
The implementation intentionally imports neither \texttt{scipy} nor
\texttt{sympy}; all symbolic handling is done through Symbolica evaluators.

\section{Implementation of the modular \texttt{FSD\_v2} prototype}

The sibling folder \texttt{FSD\_v2} is the modular reimplementation of the
prototype.  Its implementation is organized so that the black-box rule is
visible directly in the code layout:
\begin{itemize}
  \item \texttt{FSD.py} is only the command-line steering layer.  It parses
  options, validates kinematics, builds the request object, prints summaries,
  calls the benchmark, launches the integration, and renders the final table.
  \item \texttt{sectors\_generator.py} owns the declarative sector data:
  sector maps \(X_s\), regular Jacobians \(J_s\), endpoint monomials \(M_s\),
  singular axes, and generated Symbolica callbacks for these expressions.
  \item \texttt{integrand.py} owns the topology definition and the generic
  \texttt{SectorProcessor}.  The topology retains \(\U\) and \(\F\) for
  display and evaluator construction, but the processor never substitutes a
  sector map into them symbolically.
  \item \texttt{integrator.py} owns the Havana sampling loop, manual Laurent
  coefficient accumulation, progress reporting, target-accuracy stopping, and
  hot-path timing profile.
  \item \texttt{benchmark.py}, \texttt{formatting.py}, \texttt{definitions.py},
  and \texttt{utils.py} isolate OneLOopBridge access, colored output, shared
  data containers, and parenthesis uncertainty formatting.
\end{itemize}

At runtime the processor receives only a \texttt{SectorDefinition}, a
\texttt{TopologyDefinition}, and numeric sample points.  For a sector point
\(y\) it evaluates
\begin{equation}
  x = X_s(y),
  \qquad
  \hU(x),
  \qquad
  \hF(x),
  \qquad
  J_s(y),
\end{equation}
through prebuilt Symbolica evaluators.  If the sector has no endpoint
monomial, the contribution is evaluated directly.  If a monomial is declared,
the residual
\begin{equation}
  \phi_s(y)=\frac{\hF(X_s(y))}{M_s(y)}
\end{equation}
is evaluated by this quotient in the interior.  On endpoint samples the same
quantity is recovered from the Taylor coefficients of a dualized black-box
\(\hF\) evaluator.  The dual jets of the known sector map provide the chain
rule input, so no symbolic expansion of \(\F(X_s(y))\) is performed.

The generic processor then constructs
\begin{equation}
  g_s(y;\eps)
  =
  J_s(y)\,
  \hU(X_s(y))^{p_U+2\eps}\,
  \phi_s(y)^{-p_F-\eps},
\end{equation}
and applies the declared localized subtraction formula.  Zero, one, and two
singular axes are supported explicitly.  Stronger endpoint structures raise an
error instead of being guessed.

The Havana driver samples one discrete sector dimension and the common
continuous sector dimension.  Lower-support subtraction pieces are localized
into the same highest-dimensional sample space by dummy unit-density
dependence on unused variables.  Each sampled point therefore contributes to
all Laurent coefficients through the same manual accumulator,
\begin{equation}
  A_k
  =
  \frac{1}{N}\sum_{n=1}^{N}
  w_n\,I_{k,s_n}(y_n),
  \qquad
  k\in\{-2,-1,0\}.
\end{equation}
Havana is trained only on the finite-part magnitude used as an importance
function.  The adaptive grid is cloned at the beginning of an iteration; batch
training samples are accumulated into the clone and merged back into the live
grid only at the end of a completed iteration.

The target-accuracy stop is evaluated after each accumulated batch, not only
after a full Havana iteration.  Thus the final returned coefficients contain
all completed batches up to the stopping point.  In parallel mode the driver
keeps only a worker-sized window of batches in flight, so early stopping does
not submit the entire iteration unnecessarily.  The option
\(\texttt{--batch-size}\) therefore controls both vectorization size and
mid-iteration stopping granularity.

The timing profile reported by the progress bar and final table is a work-time
profile:
\begin{equation}
  T_{\rm prof}
  =
  T_{\rm Python}
  +
  T_{\rm eval}
  +
  T_{\rm Havana}.
\end{equation}
Here \(T_{\rm eval}\) is the sum of Symbolica evaluator wall times over all
workers, \(T_{\rm Python}\) is NumPy and Python glue on workers plus master-side
submission, result retrieval, accumulation, and progress rendering, and
\(T_{\rm Havana}\) is Havana sampling, training accumulation, grid merge, and
grid update time.  The displayed evaluator average is
\begin{equation}
  \langle t_{\rm eval}\rangle
  =
  \frac{10^6\,T_{\rm eval}}{N_{\rm accepted}}
  \quad
  \text{microseconds per sample per worker},
\end{equation}
because \(T_{\rm eval}\) is already summed over workers.

The current supported commands are, for example,
\begin{align}
  &\texttt{FSD\_v2/.venv/bin/python FSD.py --integral triangle --s 1.0 --m 1.0},
  \\
  &\texttt{FSD\_v2/.venv/bin/python FSD.py --integral triangle --s -1.0 --m 0.0},
  \\
  &\texttt{FSD\_v2/.venv/bin/python FSD.py --integral box --s12 0.5 --s23 0.7 --m 1.0},
  \\
  &\texttt{FSD\_v2/.venv/bin/python FSD.py --integral box --s12 -1.0 --s23 -2.0 --m 0.0}.
\end{align}
Massless timelike commands, such as
\(\texttt{--integral triangle --s 1.0 --m 0.0}\), are rejected by validation.
OneLOopBridge is required in all runs because the implementation always
compares the Laurent coefficients to a benchmark value.

\end{document}
